\Large{Adversarial Variational Optimization of Non-Differentiable Simulators}

### 2017-07-22 tex/paper.tex
contour levels
fix rajesh small comments

In parallel, with the introduction of variational
auto-encoders~\citep{DBLP:journals/corr/KingmaW13} and generative adversarial
networks~\cite{goodfellow2014generative}, there has been a vibrant research
program around implicit generative models based on neural
networks~\citep{2016arXiv161003483M}.  While these implicit generative networks
also do not admit a tractable density, they are differentiable.
Generative models based on neural networks are highly parametrized and the model
parameters have no obvious interpretation. In contrast, scientific simulators
can be thought of as highly regularized generative models as they typically have
relatively few parameters and they are endowed with some level of
interpretation. In this setting, inference on the model parameters $\bftheta$ is
often of more interest than the latent variables $\bfz$.

@@

Typically in variational inference one starts with a tractable prior
$p(\bftheta)$ and a tractable likelihood $p(\bfx |\bftheta )$, thus fixing the
posterior $p(\bftheta | \bfx )$.   After optimizing $q(\bftheta | \bfpsi)$ to
approximate the posterior, one would criticize the model through the
posterior predictive distribution as advocated by Box~\citep{box1980sampling}.
In our case, we can identify the proposal distribution with the variational posterior
$q(\bftheta | \bfpsi)$, which we effectively optimize by
directly minimizing the  divergence of the corresponding posterior predictive
distribution $\qxpsi$ and the data distribution $p_r(\bfx)$, without evaluating
the intractable likelihood $p(\bfx | \bftheta )$ or specifying a prior.
Consequently, the proposed algorithm is a form of variational inference through
model criticism~\citep{box1980sampling}.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.48\textwidth]{figures/multi.pdf}
    \caption{Multidimensional continuous data.
             ({\it Left}) Density $q(\bftheta|\bfpsi)$ at the beginning of the procedure, for a proposal distribution initialized with zero mean and unit variance.
             ({\it Right}) Density $q(\bftheta|\bfpsi)$ after adversarial variational optimization ($\gamma=10$). The proposal density correctly converges towards a distribution whose density concentrates around $\bftheta^* = (1, -1)$.
             }\label{fig:multi}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.48\textwidth]{figures/weinberg.pdf}
    \caption{Electron--positron annihilation.
    ({\it Top left}) Proposal distributions $q(\bftheta|\bfpsi)$ after adversarial variational optimization. The density of the penalized distribution ($\gamma=5$) is here highly concentrated, resulting in the green mass near $\bftheta^*$.
    ({\it Top right}) Model distributions $\qxpsi$ after training. Despite the differences between their respective proposal distributions, both models closely match the observed data.
    ({\it Bottom}) Empirical estimates of the variational upper bound $U_d$ as optimization progresses.
             }\label{fig:weinberg}
\end{figure}

>>>>

In parallel, with the introduction of variational
auto-encoders~\citep{DBLP:journals/corr/KingmaW13} and generative adversarial
networks~\cite{goodfellow2014generative}, there has been a vibrant research
program around implicit generative models based on neural
networks~\citep{2016arXiv161003483M}.  While some of these models
also do not admit a tractable density, they are all differentiable by construction.
In addition, generative models based on neural networks are highly parametrized and the model
parameters have no obvious interpretation. In contrast, scientific simulators
can be thought of as highly regularized generative models as they typically have
relatively few parameters and they are endowed with some level of
interpretation. In this setting, inference on the model parameters $\bftheta$ is
often of more interest than the latent variables $\bfz$.

@@

Typically in variational inference one starts with a tractable prior
$p(\bftheta)$ and a tractable likelihood $p(\bfx |\bftheta )$, thus fixing the
posterior $p(\bftheta | \bfx )$.   After optimizing $q(\bftheta | \bfpsi)$ to
approximate the posterior, one would criticize the model through the
posterior predictive distribution~\citep{box1980sampling,rubin1984bayesianly}.
In our case, we can identify the proposal distribution with the variational posterior
$q(\bftheta | \bfpsi)$, which we effectively optimize by
directly minimizing the  divergence of the corresponding posterior predictive
distribution $\qxpsi$ and the data distribution $p_r(\bfx)$, without evaluating
the intractable likelihood $p(\bfx | \bftheta )$ or specifying a prior.
Consequently, the proposed algorithm is a form of variational inference through
model criticism.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.48\textwidth]{figures/multi.pdf}
    \caption{Multidimensional continuous data.
             ({\it Left}) Density $q(\bftheta|\bfpsi)$ at the beginning of the procedure, for a proposal distribution initialized with zero mean and unit variance.
                          Contours correspond to parameters $\bftheta$ within $1$-$2$-$3$ Mahalanobis distance units from the mean of $q(\bftheta|\bfpsi)$.
             ({\it Right}) Density $q(\bftheta|\bfpsi)$ after adversarial variational optimization ($\gamma=10$).
                           The proposal density correctly converges towards a distribution whose density concentrates around $\bftheta^* = (1, -1)$.
             }\label{fig:multi}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.48\textwidth]{figures/weinberg.pdf}
    \caption{Electron--positron annihilation.
    ({\it Top left}) Proposal distributions $q(\bftheta|\bfpsi)$ after adversarial variational optimization.
                     Contours correspond to parameters $\bftheta$ within $1$-$2$-$3$ Mahalanobis distance units from the mean of $q(\bftheta|\bfpsi)$.
                     The density of the penalized distribution ($\gamma=5$) is here highly concentrated, resulting in the green mass near $\bftheta^*$.
    ({\it Top right}) Model distributions $\qxpsi$ after training. Despite the differences between their respective proposal distributions, both models closely match the observed data.
    ({\it Bottom}) Empirical estimates of the variational upper bound $U_d$ as optimization progresses.
             }\label{fig:weinberg}
\end{figure}
